As game industry immerses, playing games has become an important entertainment activity in our digital life. 
To improve players' experience, it is very important for game companies to understand 
player's behaviors and design user-specific features. As players may change their playing strategies 
during the game-play, understanding gamer's playing pattens over time becomes an important task. 

In this paper, we study the problem of understanding users' behaviors from their time series data obtained
during the game-play.  In particular, we focus on two tasks: 
1) what are the underlying playing states that controls user's behavior at each time period, 
and 2) what are the natural groups of players based on their state transitions (game playing strategy). 
We model user's time series data using Hidden Markov Model (HMM) \cite{hmm}, with the playing states as 
the hidden nodes and the players' statistics at each time period as the observed features. 
We fit the HMM model and find users' playing states by applying Viterbi decoding algorithm. 
We then find the groups of players using an unsupervised Dirichlet clustering method.

We experiment our method on a mobile strategy game developed by Zynga Inc. %, named {\it Empires and Allies}. 
Our method found three underlying playing states, and  the clustering results reveal three 
interesting groups that contains players with similar transition states within each group.
All these results are helpful to understanding players' behaviors and improve the game design. 


